Sparse spectral methods for partial differential equations on spherical caps

### 2020-11-04 papers/spherical-caps/spherical-caps.tex
added some more comments

This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps. More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
\begin{align*}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: ||(x,y)|| := \sqrt{x^2 + y^2} = \rho(z)\}
\end{align*}
and consider the case where $\alpha \in (-1,1)$, $\beta := 1$ and where $ \rho(z) := \sqrt{1-z^2}$.
This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.
\begin{align*}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
\end{align*}
For simplicity we focus on the case of a spherical cap, though there is a simple extension to a spherical band by taking $\beta \in (\alpha,1)$. The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper (these adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same). For clarity we note that, for all examples and plots in this paper, we take $\alpha = 0.2$.

We advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}. Equivalently, we can think of these as polynomials modulo the vanishing ideal $\{ x^2 + y^2 + z^2 = 1 \}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere. This is in contrast to more standard approaches based on mapping the geometry to a simpler one (e.g., a rectangle or disk) and using orthogonal polynomials in the mapped coordinates (e.g., a basis that is polynomial in the spherical coordinates $\varphi$ and $\theta$). The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients. Further, we remove the singular nature as $\alpha$ approaches $0$ that such a projection would give, since our new approach is applicable for all $\alpha \in (-1,1)$. On the spherical cap, the family of weights we consider are of the form
\begin{align*}
	\Wa(x,y,z) := (z - \alpha)^a, \qfor (x,y,z) \in \Omega,
\end{align*}
noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$. The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$. We define these to be orthogonalised lexicographically, that is,
\begin{align*}
	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
\end{align*}
where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - i} \: y^{i} \: z^{n-j}$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.

We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
\begin{align*}
	z &= \cos\varphi, \\
	x &= \sin\varphi \cos\theta = \rho(z) \cos\theta, \\
	y &= \sin\varphi \sin\theta = \rho(z) \sin\theta.
\end{align*}
we have
\begin{align*}
	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
\end{align*}
i.e. $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$. We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.  Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.  As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.

@@

Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + V(x,y,z)$ where $V$ is first approximated by a polynomial. A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix. This improves the complexity further to an optimal $O(p^2$), which is demonstrated in Figure \ref{fig:complexity}. This plot presents the time taken\footnote{measured using the \enquote{@belapsed} macro from the BenchmarkTools package in Julia.} to construct the operator for $\DeltaS + V(x,y,z)$, with $V(x,y,z) = \cos z$, as we increase the orthogonal polynomial degree $p$.

@@

\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
	Let $w : (\alpha,\beta) \: \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$
\begin{align*}
	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
\end{align*}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\begin{align*}
	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
\end{align*}
on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the Lebesgue measure on $\Omega$. 
\sotodoinline{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} 
\bstodoinline{I have changed it to dA and elaborated a bit more on it, but I am not sure on this so you could be correct that another term needs to be used. Your quadratic surfaces paper with Yuan says Lebesgue I think when discussing this. I have tried to keep it general here by using a measure definition, but we could omit this if its just going to cause confusion?}
\end{proposition}

@@

The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define notation that groups degree $N$ polynomials together[: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
\sotodoinline{The following is wrong as it only applies if $N = \infty$} 
\bstodoinline{What is better? To define them in this way for $N \to \infty$? Or to include the extra terms in this defn below? Or just state that the Jacobi operators approximate multiplication by x,y,z? This also relates to the next comments}
\begin{align}
	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

@@

While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the recurrence coefficient matrices necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings. To this end, we define our OP-building ordering as follows. For $n=0,1,2,\dots$:
\begin{align*}
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to 
\sotodoinline{Can the following be simplified? Seems like it repeats} 
\bstodoinline{Is it worth removing the previous? And just explaining that we can reorder the Jacobi matrices according to the OP/coefficient vector ordering? This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
\begin{align}
	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
\end{align}
where
\begin{align*}
	\tilde J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
\end{align*}
These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).

@@

\begin{figure}[t]
	\centering % <-- added
	\includegraphics[scale=0.35]{complexity-Nend=1000}
	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$. This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution. Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$. \bstodoinline{Redo with belapsed}}
	\centering
	\label{fig:complexity}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$. Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values. This indicates spectral convergence. \bstodoinline{Plots increased to N=200 for each}}
	\centering
	\label{fig:poisson}
\end{figure}
%
%\begin{figure}[t]
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
%	\centering
%	\end{subfigure}
%	\begin{subfigure}{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
%	\centering
%	\end{subfigure}
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
%	\centering
%	\end{subfigure}
%	\centering
%	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$. Left: Computed. Centre: Exact. Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
%	\centering
%	\label{fig:poissonexact}
%\end{figure}

>>>>

This paper develops sparse spectral methods for solving linear partial differential equations on certain subsets of the sphere -- specifically spherical caps. More precisely, we consider the solution of partial differential equations on the domain $\Omega$ given by
\begin{align*}
	\Omega := \{(x,y,z) \in \R^3 \quad | \quad \alpha < z < \beta, \: x^2 + y^2 + z^2 = 1\}
\end{align*}
where $\alpha \in (-1,1)$ and $\beta := 1$.
This framework then yields that $\Omega$ is a spherical cap (the region of the surface of a sphere where the $z$-coordinate ranges from $\alpha$ to $1$), i.e.

\begin{remark}
For simplicity we focus on the case of a spherical cap, though there is an extension to a spherical band by taking $\beta \in (\alpha,1)$. The methods presented here translate to the spherical band case by including the necessary adjustments to the weights and recurrence relations we present in this paper. These adjustments make the mathematics more involved, which is why they are omitted here, but the approach is the same.
\end{remark}

We advocate using a basis that is polynomial in cartesian coordinates, that is, polynomial in $x$, $y$, and $z$, and orthogonal with respect to a prescribed weight: that is, multivariate orthogonal polynomials, whose construction was considered in \cite{olver2020orthogonal}. Equivalently, we can think of these as polynomials modulo the vanishing ideal $\{ x^2 + y^2 + z^2 = 1 \}$, or simply as a linear recombination of spherical harmonics that are orthogonalised on a subset of the sphere. This is in contrast to more standard approaches based on mapping the geometry to a simpler one (e.g., a rectangle or disk) and using orthogonal polynomials in the mapped coordinates (e.g., a basis that is polynomial in the spherical coordinates $\varphi$ and $\theta$). The benefit of the new approach is that we do not need to resolve Jacobians, and thereby we can achieve sparse discretisations for partial differential operators, including those with polynomial variable coefficients. Further, we remove the singular nature as $\alpha$ approaches $0$ that such a projection would give, since our new approach is applicable for all $\alpha \in (-1,1)$. On the spherical cap, the family of weights we consider are of the form
\begin{align*}
	\Wa(x,y,z) := (z - \alpha)^a, \qfor (x,y,z) \in \Omega,
\end{align*}
noting that $\Wa(x,y,z) = 0$ for $(x,y,z) \in \partial \; \Omega$ when $a > 0$. The corresponding OPs denoted $\scopnkia(x,y,z)$, where $n$ denotes the polynomial degree, $0 \le k \le n$ and $i \in \{0, \min(1,k)\}$. We define these to be orthogonalised lexicographically, that is,
\begin{align*}
	\scopnkia(x,y,z) = C_{n,k,i} \: x^{k-i} \: y^i \: z^{n-k} + (\hbox{lower order terms})
\end{align*}
where $C_{n,k,i} \neq 0$ and \enquote{lower order terms} includes degree $n$ polynomials of the form $x^{j - i} \: y^{i} \: z^{n-j}$ where $j < k$. The precise normalization arises from their definition in terms of one-dimensional OPs in Definition~\ref{def:OPconstruction}.

We consider partial differential operators involving the spherical Laplacian (the Laplace--Beltrami operator): in spherical coordinates 
\begin{align*}
	z &= \cos\varphi, \\
	x &= \sin\varphi \cos\theta = \rho(z) \cos\theta, \\
	y &= \sin\varphi \sin\theta = \rho(z) \sin\theta.
\end{align*}
where $ \rho(z) := \sqrt{1-z^2}$, we have
\begin{align*}
	\DeltaS &= {1 \over \sinphi} \ppphi \Big( \sinphi \ppphi \Big) + {1 \over \sin^2 \varphi} \ppthetatwo = {1 \over \rho} \ppphi \Big( \rho \ppphi \Big) + {1 \over \rho^2} \ppthetatwo
\end{align*}
i.e. $\DeltaS f(\xvec) = \Delta f({\xvec \over \norm{\xvec}})$ for $\xvec := (x,y,z) \in \R^3$. We do so by considering the component operators $\rho {\partial \over \partial \varphi}$ and ${\partial \over \partial \theta}$ applied to OPs with a specific choices of weight so that their discretisation is sparse, see Theorem~\ref{theorem:sparsityofdifferentialoperators}.  Sparsity comes from expanding the domain and range of an operator using different choices of the parameter $a$, a la the ultraspherical spectral method for intervals \cite{olver2013fast}, triangles \cite{olver2019triangle} and disk-slices and trapeziums \cite{snowball2019sparse}, and the related work on sparse discretisations on disks \cite{vasil2016tensor} and spheres \cite{vasil2019tensor,lecoanet2019tensor}.  As in the  disk-slice case in 2D \cite{snowball2019sparse}, we use an integration-by-parts argument to deduce the sparsity structure.

@@

Note that we consider partial differential operators that are not necessarily rotational invariant:  for example, one can use these techniques for  Schr\"odinger operators $\DeltaS + V(x,y,z)$ where $V$ is first approximated by a polynomial. A nice feature though is that if the partial differential operator is invariant with respect to rotation around the $z$ axes (e.g., a Schr\"odinger operator with potential $V(z)$) the discretisation decouples, and can be reordered as a block-diagonal matrix. This improves the complexity further to an optimal $O(p^2$), which is demonstrated\sotodo{This isn't actually shown in the referenced figure, which shows a non-radially symmetric potential Also, we need to make $V$ vs $v$ consistent with the figures and example} in Figure \ref{fig:complexity} with $V(x,y,z) = \cos z$.

@@

\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
	Let $w : (\alpha,\beta) \: \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$\sotodo{Make notation consistent with Def 3, i.e., whether you use $\xvec$ or not}
\begin{align*}
	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
\end{align*}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\begin{align*}
	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
\end{align*}
on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
\end{proposition}

@@

The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define notation that groups degree $N$ polynomials together: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
\sotodoinline{The following is wrong as it only applies if $N = \infty$} 
\bstodoinline{What is better? To define them in this way for $N \to \infty$? Or to include the extra terms in this defn below? Or just state that the Jacobi operators approximate multiplication by x,y,z? This also relates to the next comments}
\sotodoinline{If we get rid of them as suggested below this is no longer an issue}
\begin{align}
	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

@@

While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the recurrence coefficient matrices necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings. To this end, we define our OP-building ordering as follows. For $n=0,1,2,\dots$:
\begin{align*}
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to 
\sotodoinline{Can the following be simplified? Seems like it repeats} 
\bstodoinline{Is it worth removing the previous? And just explaining that we can reorder the Jacobi matrices according to the OP/coefficient vector ordering? This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
\sotodoinline{I agree: get rid of the previous definition and remove the tilde from these polynomials, as they are the "default" ones}
\begin{align}
	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
\end{align}
where
\begin{align*}
	\tilde J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
\end{align*}
These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).

@@

\begin{figure}[t]
	\centering % <-- added
	\includegraphics[scale=0.35]{complexity-Nend=1000}
	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$. This shows that the time to build and solve the system $\big[\DeltaS + V(x,y,z) \big] \: u(x,y,z) = f(x,y,z)$ is roughly of order $\bigO(N^2)$, where $N$ is the degree to which we approximate the solution. Here, we used $f = -2e^{x}yz(2+x) + (z - \alpha) e^{x} (y^3 + z^2 y - 4xy - 2y)$. \bstodoinline{Redo with belapsed}}
	\centering
	\label{fig:complexity}
\end{figure}

\begin{figure}[t]
	\begin{subfigure}{0.5\textwidth}
	\includegraphics[scale=0.25]{poisson-u=wyexpx-N=60}
	\centering
	%\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.3]{solutionblocknorms-poisson-epsilonfun-N=200}
        	%\label{fig:solutionblocknorms-poisson}
	\end{subfigure}
	\caption{Left: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y,z) = - 2 e^x y z (2+x) + \genjacw^{(1,0)}(z) e^x (y^3 + z^2 y - 4xy - 2y)$. Right: The norms of each block of the computed solution of the Poisson equation with right hand side function $f(x,y,z) = \norm{\xvec - (\epsilon + 1 / \sqrt{3}) \; (1,1,1)^\top }$ for different $\epsilon$ values. This indicates spectral convergence.}
	\centering
	\label{fig:poisson}
\end{figure}
%
%\begin{figure}[t]
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx}
%	\centering
%	\end{subfigure}
%	\begin{subfigure}{0.3\textwidth}
%	\centering
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-actual}
%	\centering
%	\end{subfigure}
%	\begin{subfigure}{0.3\textwidth}
%	\includegraphics[scale=0.3]{solution-exact-poisson-diskslice-alpha=0p2-beta=0p8-u=y3expx-compare}
%	\centering
%	\end{subfigure}
%	\centering
%	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions compared with the exact solution $u(x,y) = \Wiii (x,y) y^3 \exp(x)$. Left: Computed. Centre: Exact. Right: Plot of the error (colourbar is shown to demonstrate magnitude of the error is of the order $10^{-17}$)}
%	\centering
%	\label{fig:poissonexact}
%\end{figure}
